---
title: "Calculating Life Table Values"
---

```{r bio239 line 5, include = FALSE, echo = FALSE}
require(tidyverse)
require(knitr)
```

<div style = "margin-bottom:75px;">
</div>

<div style = "margin-bottom:75px;">


For species that reproduce at **discrete times** during the year, such as birds that lay eggs in spring, the **geometric growth model** describes changes in population size at regular intervals. Population growth in these species is measured by comparing the population size at a certain time of year to the population size at the same time the preceding year. The **geometric growth model** specifies this ratio using the Greek symbol lambda, $\lambda$.

$N_{t} = N_{0}\lambda^{t}$

</div>

<div style = "margin-bottom:75px;">
**Version A: Population Growth in Rabbits**

Suppose that a scientist collected data on the number of rabbits in a population each year. The data are provided in the table.
</div>

<div style = "margin-bottom:50px;">

Gray Squirrel Life Table

```{r table1, echo = FALSE, message = FALSE, warnings = FALSE, results ='asis'}
tabl <- "
|(x) |(n_{x})|(l_{x})|(b_{x})|(l_{x}) x (b_{x})|(x) x (l_{x}) x (b_{x})|
|:--:|:-----:|:-----:|:-----:|:---------------:|:---------------------:|
| 0  | 530   | 1.000 | 0.05  |                 |                       |
| 1  | 134   | 0.253 | 1.28  |                 |                       |
| 2  | 56    | 0.116 | 2.28  |                 |                       |
| 3  | 39    | 0.089 | 2.28  |                 |                       |
| 4  | 23    | 0.058 | 2.28  |                 |                       |
| 5  | 12    | 0.039 | 2.28  |                 |                       |
| 6  | 5     | 0.025 | 2.28  |                 |                       |
| 7  | 2     | 0.022 | 2.28  |                 |                       |
"
cat(tabl)
```
</div>