---
title: "Frequency-Dependent Selection"
---

```{r bio239 line 5, include = FALSE, echo = FALSE}
require(tidyverse)
```

<div style = "margin-bottom:25px;">
From page 195 in the textbook
</div>

<div style = "margin-bottom:25px;">
We can better understand frequency-dependent selection if we work through an example with real numbers. The American black vulture (*Coragyps atratus*) is a large bird that feeds on animal carcasses throughout much of North and South America. The female vulture typically lays two eggs in the nest.

If a population is composed of five females and two males and we assume that a male can breed with more than one female, how many copies of genes, on average, does each male and female contribute to the next generation?
</div>

<div style = "margin-bottom:25px;">
**Total number of eggs** = 10  
**Average female fitness** = 10 female gene copies / 5 females = 2 gene copies / female  
**Average male fitness** =   
</div>

<div style = "margin-bottom:25px;">
1. Using the assumptions above, calculate the number of gene copies per male and per female under the following two scenarios:
a. Four males and five females
b. Six males and five females

2. Given the four scenarios that we have explored, what sex ratio will be favored by natural selection over the long term?
</div>

<div style = "margin-bottom:25px;">
Scatterplot in R
</div>

<div style = "margin-bottom:25px;">
```{r lesson3 23, include = TRUE}
scatter.df <- data.frame(
  
  x = c(1, 2, 3, 4),
  
  y = c(4, 3, 2, 1),
          
  number = c("a", "b", "c", "d")
          
)

scatter.df

scatter.gg <- ggplot(data = scatter.df, aes(x = x, y = y)) +
  geom_point(aes(color = number), size = 3, alpha = 0.6)
scatter.gg
```
</div>        