---
title: "Population Growth and Regulation"
---

```{r bio239 line 5, include = FALSE, echo = FALSE}
require(tidyverse)
require(knitr)
```

<div style = "margin-bottom:75px;">
</div>

<div style = "margin-bottom:75px;">

Different models of population growth account for different modes of reproduction. Species that reproduce continuously throughout a given period of time, such as humans, can grow **exponentially**. The **exponential growth model** calculates the population size at a given time ($N_{t}$) by multiplying the starting population size ($N_{0}$) by an exponential term. This model raises $e$, the base of the natural log, by the product of the growth rate ($r$) and the interval of time ($t$).

$N_{t} = N_{0}e^{rt}$

For species that reproduce at **discrete times** during the year, such as birds that lay eggs in spring, the **geometric growth model** describes changes in population size at regular intervals. Population growth in these species is measured by comparing the population size at a certain time of year to the population size at the same time the preceding year. The **geometric growth model** specifies this ratio using the Greek symbol lambda, $\lambda$.

$N_{t} = N_{0}\lambda^{t}$

</div>

<div style = "margin-bottom:75px;">
**Version A: Population Growth in Rabbits**

Suppose that a scientist collected data on the number of rabbits in a population each year. The data are provided in the table.
</div>

<div style = "margin-bottom:50px;">
```{r table1, echo = FALSE, message = FALSE, warnings = FALSE, results ='asis'}
tabl <- "
|Year|Population Size| 
|:--:|:--:|
| 1 | 10  |              
| 2 | 20  |              
| 3 | 40  |            
| 4 | 80  |            
| 5 | 160 |    
| 6 | 320 | 
"
cat(tabl)
```
</div>

<div style = "margin-bottom:50px;">
**Question 1**
Using these data, graph the number of individuals by year. 
</div>

<div style = "margin-bottom:50px;">
```{r bio239 line 53, echo = TRUE}
t <- c ( 0, 1, 2, 3, 4, 5 )
N <- c ( 10, 20, 40, 80, 160, 320 )
qplot( x = t, y = N, geom = c ( "line", "point" ) )
```
</div>

<div style = "margin-bottom:50px;">
**Question 2**
From the data, calculate $\lambda$ with the equation $\lambda = N_{t}+1 / N_{t}$, where $N_{t}$ equals the population size at one time point and Nt+1 equals the population size one year later.
</div>

<div style = "margin-bottom:50px;">
**Question 3**
Predict the population size, $N_{10}$, at year 10 using the equation $N_{t}  = N_{0}/\lambda^{t}$, where $N_{0}$ is the population size at year 1 and *t* is year.
</div>

<div style = "margin-bottom:50px;">
**Question 4**
Compare the population sizes you calculated at year 10. How is population growth affected by $\lambda$?
</div>

<div style = "margin-bottom:50px;">
**Question 5**
Why is it more appropriate to use a geometric model of growth than an exponential model for these data?
</div>

<div style = "margin-bottom:75px;">
**Version B: Population Growth in White-Tailed Deer**

Suppose that a scientist collected data on the number of white-tailed deer in a population each year. The data are provided in the table.
</div>

<div style = "margin-bottom:50px;">
```{r table2, echo = FALSE, message = FALSE, warnings = FALSE, results ='asis'}
tabl <- "
|Year|Population Size| 
|:--:|:--:|
| 1 | 10  |              
| 2 | 17  |              
| 3 | 29  |            
| 4 | 49  |            
| 5 | 84 |    
| 6 | 142 | 
"
cat(tabl)
```
</div>

<div style = "margin-bottom:50px;">
**Question 1**
Using these data, graph the number of individuals by year. 
</div>

<div style = "margin-bottom:50px;">
```{r bio239 line 108, echo = TRUE}
t <- c ( 0, 1, 2, 3, 4, 5 )
N <- c ( 10, 17, 29, 49, 84, 142 )
qplot( x = t, y = N, geom = c ( "line", "point" ) )
```
</div>

<div style = "margin-bottom:50px;">
**Question 2**
From the data, calculate $\lambda$ with the equation $\lambda = N_{t}+1 / N_{t}$, where $N_{t}$ equals the population size at one time point and Nt+1 equals the population size one year later.
</div>

<div style = "margin-bottom:50px;">
**Question 3**
Predict the population size, $N_{10}$, at year 10 using the equation $N_{t}  = N_{0}/\lambda^{t}$, where $N_{0}$ is the population size at year 1 and *t* is year.
</div>

<div style = "margin-bottom:50px;">
**Question 4**
Compare the population sizes you calculated at year 10. How is population growth affected by $\lambda$?
</div>

<div style = "margin-bottom:50px;">
**Question 5**
Why is it more appropriate to use a geometric model of growth than an exponential model for these data?
</div>

<div style = "margin-bottom:75px;">
**Version C: Population Growth in Black Bears**

Suppose that a scientist collected data on the number of black bears in a population each year. The data are provided in the table.
</div>

<div style = "margin-bottom:50px;">
```{r table3, echo = FALSE, message = FALSE, warnings = FALSE, results ='asis'}
tabl <- "
|Year|Population Size| 
|:--:|:--:|
| 1 | 10  |              
| 2 | 13  |              
| 3 | 17  |            
| 4 | 22  |            
| 5 | 29 |    
| 6 | 37 | 
"
cat(tabl)
```
</div>

<div style = "margin-bottom:50px;">
**Question 1**
Using these data, graph the number of individuals by year. 
</div>

<div style = "margin-bottom:50px;">
```{r bio239 line 163, echo = TRUE}
t <- c ( 0, 1, 2, 3, 4, 5 )
N <- c ( 10, 13, 17, 22, 29, 37 )
qplot( x = t, y = N, geom = c ( "line", "point" ) )
```
</div>

<div style = "margin-bottom:50px;">
**Question 2**
From the data, calculate $\lambda$ with the equation $\lambda = N_{t}+1 / N_{t}$, where $N_{t}$ equals the population size at one time point and Nt+1 equals the population size one year later.
</div>

<div style = "margin-bottom:50px;">
**Question 3**
Predict the population size, $N_{10}$, at year 10 using the equation $N_{t}  = N_{0}/\lambda^{t}$, where $N_{0}$ is the population size at year 1 and *t* is year.
</div>

<div style = "margin-bottom:50px;">
**Question 4**
Compare the population sizes you calculated at year 10. How is population growth affected by $\lambda$?
</div>

<div style = "margin-bottom:50px;">
**Question 5**
Why is it more appropriate to use a geometric model of growth than an exponential model for these data?
</div>


