---
title: "Population Growth and Regulation"
---

```{r bio239 line 5, include = FALSE, echo = FALSE}
require(tidyverse)
require(knitr)
```

<div style = "margin-bottom:75px;">
</div>

<div style = "margin-bottom:75px;">

Version A: Population Growth in Rabbits

Suppose that a scientist collected data on the number of rabbits in a population each year. The data are provided in the table.

</div>

<div style = "margin-bottom:50px;">
```{r table1, echo = FALSE, message = FALSE, warnings = FALSE, results ='asis'}
tabl <- "
|Year|Population Size| 
|:--:|:--:|
| 1 | 10  |              
| 2 | 20  |              
| 3 | 40  |            
| 4 | 80  |            
| 5 | 160 |    
| 6 | 320 | 
"
cat(tabl)
```
</div>

<div style = "margin-bottom:50px;">
```{r bio239 line 43, echo = TRUE}
t <- c ( 0, 1, 2, 3 )
N <- c ( 2, 4, 8, 16 )
qplot( x = t, y = N, geom = c ( "line", "point" ) )
```
</div>

<div style = "margin-bottom:50px;">
**Question 1**
Using these data, graph the number of individuals by year. 
</div>

<div style = "margin-bottom:50px;">
**Question 2**
From the data, calculate $\lambda$ with the equation $\lambda = N_{t}+1 / N_{t}$, where $N_{t}$ equals the population size at one time point and Nt+1 equals the population size one year later.
</div>

<div style = "margin-bottom:50px;">
**Question 3**
Predict the population size, $N_{10}$, at year 10 using the equation $N_{t}  = N_{0}/\lambda^{t}$, where $N_{0}$ is the population size at year 1 and *t* is year.
</div>

<div style = "margin-bottom:50px;">
**Question 4**
Compare the population sizes you calculated at year 10. How is population growth affected by $\lambda$?
</div>

<div style = "margin-bottom:50px;">
**Question 5**
Why is it more appropriate to use a geometric model of growth than an exponential model for these data?
</div>






